---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
library(vegawidget)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  dev = "svg"
)
```

# liminal

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

The goal of `liminal` is to provide diganostics and visual analytics for 
understanding embedding algorithms such as tSNE. 
It has been [shown](https://distill.pub/2016/misread-tsne/) 
that in order to produce an effective embedding may
one have to play with hyperparamters and various settings for these algorithms.

`liminal` aims to increase the effective use of these algorithms by
combining them with interactive and dynamic 
graphics (via the vegawidget and shiny packages) and with a technique
from multivariate stastics called the tour (via the tourr package). 
Briefly, a tour is a sequence of interploated projections of multivariate data onto
lower dimensional space. The sequence is displayed as a dynamic visualisation,
and enables us to see the shadows of the high dimensional data makes in a
lower dimensional view. By combining the tour with embedding algorithms,
we can see the following:

1. whether distances in the embedding view are meaningful
2. the local and global structure of the data
3. identify 'interesting' shapes or points in the data

## Quick start

Let's take a look at a high-dimensional
tree structured data set:

```{r init}
library(liminal)
dim(fake_trees)
```

This data set has 3000 observations and 100 numeric variables,
and has ten branches. We can generate principal components:

```{r pc-view}
pcs  <- prcomp(fake_trees[, -ncol(fake_trees)])
# var explained
head(cumsum(pcs$sdev / sum(pcs$sdev)))
```

And visualise the results as interactive scatter plot by augmenting the orginal data

```{r pca-xy}
fake_trees <- dplyr::bind_cols(fake_trees, as.data.frame(pcs$x))
limn_xy(fake_trees, x = PC1, y = PC2, color = branches)
```

We see some separation of the branches in first two principal components,
however, we can't see each of the branches clearly or their relation to each
other. We can tourr them to get an idea of underlying structure:

```{r, eval = FALSE}
# this loads a shiny app
limn_tour(fake_trees, PC1:PC15, color = branches)
```

The blue branch is hidden and connects all other branches. 

We can also compare this to t-SNE embedding:

```{r tsne-xy}
set.seed(2099)
tsne <- Rtsne::Rtsne(dplyr::select(fake_trees, dplyr::starts_with("dim")))

tsne_df <- data.frame(tsneX = tsne$Y[,1],
                      tsneY = tsne$Y[,2],
                      branches = fake_trees$branches)
limn_xy(tsne_df, x = tsneX, y = tsneY, color = branches)
```

The topology is a little messed up. We can see where our embedding is different
via a linked scatter plot:

```{r, eval = FALSE}
limn_tour_xylink(dplyr::select(fake_trees, PC1:PC10, branches),
                 tsne_df,
                 x_color = branches, 
                 y_color = branches)
```
