---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# liminal

<!-- badges: start -->
[![R build status](https://github.com/sa-lee/liminal/workflows/R-CMD-check/badge.svg)](https://github.com/sa-lee/liminal/actions)
<!-- badges: end -->

The goal of **liminal** is to provide complementary visualisations for use with
understanding embedding algorithms such as tSNE.  It has been
[shown](https://distill.pub/2016/misread-tsne/) that in order to produce an
'effective' embedding one may have to play with hyperparamters and various
settings for these algorithms.

See the [liminal vignette](https://sa-lee.github.io/liminal/articles/liminal.html) 
for details of package usage.

## Quick Start

The development version of **liminal** can be installed as follows:

```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("sa-lee/liminal")
```


## About

**liminal** aims to assist with the use of embedding algorithms by combining
them with interactive and dynamic graphics (via the **vegawidget** and **shiny**
packages) and with a technique from multivariate statistics called the tour (via
the **tourr** package).  Briefly, a tour is a sequence of interpolated projections
of multivariate data onto lower dimensional space. The sequence is displayed as
a dynamic visualisation, and enables us to see the shadows the high dimensional
data makes in a lower dimensional view. By combining the tour with embedding
algorithms, we can see the following:

1. whether distances in the embedding view are meaningful 
2. the local and global structure of the data 
3. identify 'interesting' shapes or points in the data

The **liminal** package easily incorporates into a exploratory data analysis 
workflow, and tracks user events such as brushes for downstream analysis.




