---
title: "liminal"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{liminal}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(ggplot2)
theme_set(theme_bw())
```


Let's take a look at a high-dimensional
tree structured data set:

```{r init}
library(liminal)
dim(fake_trees)
```

This data set has 3000 observations and 100 numeric variables,
and has ten branches. We can generate principal components:

```{r pc-view}
pcs  <- prcomp(fake_trees[, -ncol(fake_trees)])
# var explained
head(cumsum(pcs$sdev / sum(pcs$sdev)))
```

And visualise the results as scatter plot by augmenting the orginal data

```{r pca-xy}
fake_trees <- dplyr::bind_cols(fake_trees, as.data.frame(pcs$x))
ggplot(fake_trees, aes(x = PC1, y = PC2, color = branches)) +
  geom_point() +
  scale_color_manual(values = limn_pal_tableau10())
```

We see some separation of the branches in first two principal components,
however, we can't see each of the branches clearly or their relation to each
other. 

We can tour the components that represent most of the variation in the 
data to get an idea of underlying structure. The code below generates a dynamic scatter plot by touring the first fifteen principal components. 

The following code generates a shiny application, the provides an interface to the tour:

```{r, eval = FALSE}
# this loads a shiny app
limn_tour(fake_trees, PC1:PC15, color = branches)
```
The interace consists of the tour view which is a dynamic scatterplot
and an axis view which corresponds to magnitude and direction of 
the generated basis vectors. 

Brushing on the tour view is activated with the shift key plus a mouse drag. It will highlight points that fall inside the brush and
pause the current view. 

There are several additional interactions available on this view:
  * There is a play button, that when pressed will start the tour.
  * There is also a text view of the half range which is the maximum squared
    Euclidean distance between points in the tour view. The half range
    is a scale factor for projections and can be thought of as a way
    of zooming in and out on points. It can be dynamically modified by scrolling
    (via a mouse-wheel). To reset double click the tour view.
  * The legend can be toggled to highlight groups of points with
    shift+mouse-click. Multiple groups can be selected in this way. To
    reset double click the legend title.

From tour view, we can see that the blue branch is hidden and forms the
backbone of the tree. 

We can also compare this to t-SNE embedding run with default settings:

```{r tsne-xy}
set.seed(2099)
tsne <- Rtsne::Rtsne(dplyr::select(fake_trees, dplyr::starts_with("dim")))

tsne_df <- data.frame(tsneX = tsne$Y[,1],
                      tsneY = tsne$Y[,2],
                      branches = fake_trees$branches)
ggplot(tsne_df, aes(x = tsneX, y = tsneY, color = branches)) +
  geom_point() +
  scale_color_manual(values = limn_pal_tableau10())
```

The topology is a little messed up as the blue branch is now broken into two 
distinct pieces. 

We can see where our embedding is different via a linked tour:

```{r, eval = FALSE}
limn_tour_xylink(dplyr::select(fake_trees, PC1:PC10, branches),
                 tsne_df,
                 x_color = branches, 
                 y_color = branches)
```
 This function requires two tables that will be linked together in seperate views. The tour interface is the same as above, except now
 brushing on the tour view will highlight points on the right hand side
 scatter plot. The right side scatterplot view is an interactive scatterplot. Brushing on the right view is activated via click and drag movements. It will highlight points in the tour view but the type of highlighting depends on the brush mode selected, the default is simple
 linked brushing. There are two other brushing modes that can be modified via radio buttons. Both neighbors and distance based brushing operate on the k-NN graph obtained from the control panel.
