---
title: "pbmc3k"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{pbmc3k}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Here we explore the use of `sneezy` on the 10x genomics PBMC3k single-cell
dataset. We follow the preprocessing and quality control steps from
[Chapter 25 of Orchestrating Single Cell Analysis with R](https://osca.bioconductor.org/pbmc-3k-10x-dataset-filtered.html), and
use that to explore the linear and non-linear embedding spaces via tours
and diagnostics.

## Setup

First, we load up the raw data from the `TENxPBMCData` package,
and the analysis packages `scater` and `scran`:

```{r setup}
library(TENxPBMCData)
library(scater)
library(scran)

pbmc3k <- TENxPBMCData("pbmc3k")
pbmc3k
```

Next we follow the simple QC steps from the aforementioned vignette:

### Quality control

```{r qc}
is.mito <- grep("MT", rowData(pbmc3k)$Symbol_TENx)
stats <- perCellQCMetrics(pbmc3k, subsets=list(Mito=is.mito))
high.mito <- isOutlier(stats$subsets_Mito_percent, nmads=3, type="higher")
pbmc3k <- pbmc3k[,!high.mito]
```

### Normalization

```{r}
pbmc3k <- logNormCounts(pbmc3k)
```

### Variance modelling

```{r}
dec3k <- modelGeneVar(pbmc3k)
chosen.hvgs <- which(dec3k$bio > 0)
```


## The TourExperiment object and a first embedding

We provide a simple interface to constructing a `TourExperiment` from
a SingleCellExperiment object. Once we have this object, we can construct
linear embedding using PCA. Since, the underlying object is backed by an HDF5 
file we use randomized PCA for speed purposes.

```{r}
library(sneezy)
set.seed(2099)
pbmc3k_te <- TourExperiment(pbmc3k)
pbmc3k_te <- embed_linear(pbmc3k_te, 
                          num_comp = 10, 
                          .on = "logcounts", 
                          .subset = chosen.hvgs, 
                          .engine = pca_random())
```

The arguments above add an embedding to the pbmc3k_te object,
using randomized PCA, only on the highly variable genes we've constructed 
above. Since, the procedure is random, we also set a seed. 


```{r}

```


The usual approach here is to inspect the results of a cluster
analysis on top of an embedding, here we have used the standard 


```{r}
g <- buildSNNGraph(pbmc3k_te, k=15, use.dimred = 'pca_random')
clust <- igraph::cluster_louvain(g)
pbmc3k_te$cluster <- factor(igraph::membership(clust))
```

We can also estimate a basis set via a grand tour: 

```{r}
pbmc3k_te <- generate_bases(
  pbmc3k_te,
  .on = "pca_random",
  max_bases = 10,
  clamp  = FALSE,
  .engine = tourr::grand_tour(),
)
pbmc3k_te
```

This creates an array with dimensions `r dim(basisSet(pbmc3k_te))`
named `pca_random`. To access the underlying array use 
`basisSet(pbmc3k_te)`.

We can then view the principal components using the `view_tour_xy()`, this is a shiny application. You can view an example of it here:

```{r, eval = FALSE}
sneezy_tour(pbmc3k_te)
```

