---
title: "pdfsense: or understanding high dimensional parameter space"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{pdfsense}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


This example is modified from the paper by Laa et al., here we show how we can
use a tour to explore principal components space and any non-linear structure and clusters via non-linear embeddings.

## Setting up the data 

Data were obtained from CT14HERA2 parton distribution function
fits as used in Laa et al., 2018. There are 28 directions in the parameter
space of parton distribution function fit, each point in the variables
labelled X1-X56 indicate moving +- 1 standard devation from the 'best' 
(maximum likelihood estimate) fit of the function. Each observation has
all predictions of the corresponding measurement from an experiment. 
 (see table 3 in that paper for more explicit details). 
 
The remaining columns are:
 
 * InFit: A flag indicating whether an observation entered the fit of
   CT14HERA2 parton distribution function
 * Type: First number of ID
 * ID: contains the identifier of experiment, 1XX/2XX/5XX correpsonds
 to Deep Inelastic Scattering (DIS) / Vector Boson Production (VBP) / 
  Strong Interaction (JET). Every ID points to an experimental paper.   
 * pt: the per experiment observational id
 * x,mu: the kinematics of a parton. x is the parton momentum fraction, and
 mu is the factorisation scale.

First, we take the load the data as a data.frame:

```{r pdfsense-prepare}
library(liminal)
data(pdfsense)
```

## Linear embeddings and the tour

First we can estimate all `nrow(pdfsense)` principal components
using on the parton distribution fits:
```{r pdfsense}
pcs  <- prcomp(pdfsense[, 7:ncol(pdfsense)])
```

Using this data structure, we can produce a screeplot:
```{r, echo = TRUE}
res <- data.frame(
  component = 1:56, 
  variance_explained = cumsum(pcs$sdev / sum(pcs$sdev))
)

limn_xy(res, x = component, y = variance_explained)
```

Approximately %70 of the variance in the pdf fits are explained by the first 15 principal components.

Next we augment our original data with the principal components:

```{r}
pdfsense <- dplyr::bind_cols(
  pdfsense, 
  as.data.frame(pcs$x)
)
pdfsense$Type <- factor(pdfsense$Type)
```

We can view a simple tour via`limn_tour()` and color points
by their experimental group

```{r, eval = FALSE}
limn_tour(pdfsense, PC1:PC6, Type)
```


## Non-Linear embeddings

Now we can set up a non-linear embedding via t-SNE, here
we embed all 56 principal components.
```{r}
set.seed(3099)
tsne <- Rtsne::Rtsne(dplyr::select(pdfsense, PC1:PC56),
                     pca = FALSE, 
                     normalize = TRUE,
                     perplexity = 50,
                     exaggeration_factor = nrow(pdfsense) / 10)
```

Again we can augment the embedding onto the pdfsense data:

```{r tsne}
tsne_embedding <- as.data.frame(tsne$Y)
tsne_embedding <- dplyr::rename(tsne_embedding,
                                tsneX = V1, tsneY = V2)
tsne_embedding$Type <- pdfsense$Type
```

We can view the clusters:

```{r}
limn_xy(tsne_embedding, tsneX, tsneY, Type,
        brush_transform = "median")
```


We can link a tour view next to the embedding to give us
a clear picture of the clustering:

```{r, eval = FALSE}
limn_tour_xylink(x = dplyr::select(pdfsense, PC1:PC6, Type),
                 y = tsne_embedding,
                 x_color = Type,
                 y_color= Type)
```



