---
title: "pdfsense: or understanding high dimensional parameter space"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{pdfsense}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


This example is modified from the paper by Laa et al., here we show how we can
use a tour to explore principal components space and any non-linear structure
and clusters via non-linear embeddings.

## Setting up the data 

Data were obtained from CT14HERA2 parton distribution function
fits as used in Laa et al., 2018. There are 28 directions in the parameter
space of parton distribution function fit, each point in the variables
labelled X1-X56 indicate moving +- 1 standard devation from the 'best' 
(maximum likelihood estimate) fit of the function. Each observation has
all predictions of the corresponding measurement from an experiment. 
 (see table 3 in that paper for more explicit details). 
 
The remaining columns are:
 
 * InFit: A flag indicating whether an observation entered the fit of 
   CT14HERA2 parton distribution function
 * Type: First number of ID
 * ID: contains the identifier of experiment, 1XX/2XX/5XX correpsonds
 to Deep Inelastic Scattering (DIS) / Vector Boson Production (VBP) / 
  Strong Interaction (JET). Every ID points to an experimental paper.   
 * pt: the per experiment observational id
 * x,mu: the kinematics of a parton. x is the parton momentum fraction, and
 mu is the factorisation scale.

First, we take the data from a data.frame to a TourExperiment data
structure:

```{r pdfsense-prepare}
library(SingleCellExperiment)
library(sneezy)
pdfsense <- TourExperiment(pdfsense, X1:X56)
pdfsense
```

## Linear embeddings and the tour

First we can estimate all `nrow(pdfsense)` principal components:
```{r pdfsense}
pdfsense <- embed_linear(pdfsense, 
                         num_comp = nrow(pdfsense),
                         center = TRUE)
pdfsense 
```

If we look at the data structure returned, we get a data structure
called a LinearEmbeddingMatrix - which holds all the loadings and components.
```{r}
pcs <- reducedDim(pdfsense, "pca_exact")
pcs
```
Using this data structure, we can produce a screeplot:
```{r, echo = TRUE}
factorData(pcs) <- cbind(
  factorData(pcs),
  variance_explained = factorData(pcs)$sdev / sum(factorData(pcs)$sdev),
  component = 1: nrow(factorData(pcs)),
  cumulative_var = cumsum(factorData(pcs)$sdev / sum(factorData(pcs)$sdev))
)

library(ggplot2)
ggplot(as.data.frame(factorData(pcs)),
       aes(x = component, y = variance_explained)) + 
  geom_point() +
  labs(x = "Principal Component", y = "Proportion Variance Explained")

```

Approximately %70 of the variance in the pdf fits are explained by the first
15 principal components.

Next we can generate a set of bases to tour on the principal components.
Note that we restrict our view to the first 6 principal components and 
generate 100 new bases via the grand tour (we choose 6 to follow the original paper):

```{r}
pdfsense <- generate_bases(pdfsense,
                           .on = "pca_exact",
                           subset = 1:6)

dim(basisSet(pdfsense))
```

We can view a simple tour via `sneezy_tour()`

```{r}
sneezy_tour(pdfsense)
```

Alternatively, we can highlight an a grouping of interest, in
this case the JET experiments:

```{r}
highlight <- 
  sneezy_tour(pdfsense, 
              col = c("grey50", "darkred")[(colData(pdfsense)$Type == 5) + 1])
highlight
```



## Non-Linear embeddings

Now we can set up a non-linear embedding via t-SNE, here
we use an approximate fit using a default perpelxity of 30.
```{r}
set.seed(1999)
pdfsense <- embed_nonlinear(pdfsense, 2, .on = "pca_exact")
tsne <- reducedDim(pdfsense, "tsne_approx")
```

And we can use ggplot2 to construct a view of the embedding:

```{r tsne}
pl <- ggplot(
  data.frame(
    sampleFactors(tsne),
    colData(pdfsense)
  )) +
  geom_point(aes(x = Dim1, y = Dim2, col = factor(Type))) +
  scale_color_manual(values = c("blue", "orange", "darkred")) +
  labs(color = "Type") +
  theme_minimal() +
  theme(axis.text.x = element_blank(), axis.text.y = element_blank()) 
```

We can link a tour view next to the embedding to give us
a clear picture of the clustering:

```{r}
view <- link_left(highlight, pl)
view 
```

Next we can look at some diagnostics for the embedding, we can view
the cluster centroids via a tour:

```{r}
pdfsense <- estimate_neighbors(pdfsense, 30, .on = "tsne_approx")
pl +
  overlay_snn_centroids(sampleFactors(tsne)[,1],
                        sampleFactors(tsne)[,2],
                        neighborSet(pdfsense),
                        size = 5)
```

```{r sneezy-centroids}
sneezy_centroids(pdfsense)
```


